Adaptive Hybrid Screening for Efficient Lasso Optimization

### 2021-09-08 jscs/response/response.tex

Thank you very much for your detailed reading of the paper and your insightful comments and questions. You raise a number of good points and we have modified the manuscript in several places based upon your feedback.  In our opinion, these revisions have nicely improved upon the original submission.

\begin{enumerate}

\item \emph{Numerical studies only focus on the computational performances of the algorithms. One may be interested in some details in the proposed algorithm: the size of variables remaining from the safe or strong rules, the number of KKT condition checks, the adaptive batch sizes, etc. It would be good to add the numerical studies for details in the algorithm.}

  This is an excellent point. We felt that some readers would definitely want to know this additional information, while others might just care that the algorithm is faster. So, in the revised manuscript we have created an appendix that presents these additional details, which do indeed lend insight into how all of the algorithms are working and how the proposed method manages to achieve efficiency gains.

@@

\item \emph{The paper assumes that the predictive variables are standardized for linear and logistic regression models. For group lasso, the additional orthogonal assumption are required for the proposed method. I’m just wondering if the proposed method is applicable without the standardization assumptions. Specifically, is it possible to develop (or exist) the sequential safe and strong rules without the standardization assumptions.}

\item \emph{In page 10, line 13, ``We also assume that post-convergence check will usually pass in the first iterations''. I’m not sure it is reasonable.}

  We now provide both a citation and our own investigations (in the appendix) to support this assumption.

@@

\item \emph{In page 8, line 19, ``the reference solutions are continually updated''. The reference solutions are not always updated. So, I’m not sure the term ``continually'' is appropriate.}

  Very good point. In the revised manuscript, we have changed this to ``the reference gets updated periodically'', which is a more accurate description.

@@

\item \emph{In Figure 3, the green line (HSSR) pattern is quite different with others. Could you explain it?}

\item \emph{In Section 6(Conclusion), there does not exist the sequential safe rules for sparse Cox regression, Poisson regression, and support vector machines. It would be good to explain the existence of the sequential strong rules for each problem.}

\item \emph{In Theorem 4.1, $x_* \in \{ x_j : \hat{\beta}_j(\lambda_l) \ne 0 \}$. Is the $x_*$ arbitrary? It seems that further explanation for $x_*$ is needed.}

\textbf{Minor corrections:}

\item \emph{In Theorem 4.1, $\lambda_0 := \max_j |x^T_j y/n|$ may be incorrect. It would be the maximum gradient value at $\hat{\beta} = 0$ of the negative log-likelihood in logistic regression model.}

\item \emph{In page 10, line 20, $|S_{l=b}|$ should be $|S_{l+b}|$.}

\item \emph{In Figure 1, the green line is not continuous. It would be better to modify Figure 1.}

\item \emph{In page 9, line 50, the ’line 12’ is better than ’line 11’ in Algorithm 1.}

\item \emph{In Theorem 4.2, $\lambda_0 := \max_j \lVert \ldots \rVert$. The $j$ should be $g$.}

### 2021-09-09 jscs/response/response.tex

\item \emph{Numerical studies only focus on the computational performances of the algorithms. One may be interested in some details in the proposed algorithm: the size of variables remaining from the safe or strong rules, the number of KKT condition checks, the adaptive batch sizes, etc. It would be good to add the numerical studies for details in the algorithm.}

  This is an excellent point. We felt that some readers would definitely want to know this additional information, while others might just care that the algorithm is faster. So, in the revised manuscript we have created an appendix that presents these additional details, which do indeed lend insight into how all of the algorithms are working and how the proposed method manages to achieve efficiency gains.

\item \emph{The paper assumes that the predictive variables are standardized for linear and logistic regression models. For group lasso, the additional orthogonal assumption are required for the proposed method. I’m just wondering if the proposed method is applicable without the standardization assumptions. Specifically, is it possible to develop (or exist) the sequential safe and strong rules without the standardization assumptions.}

    Most of the sequential safe or strong rules does not require standardization, including the rules mentioned in our manuscript. As a result, our method does not require standardization and we introduce the standardization assumption mainly for simplicity. We also add this clarification to the manuscript in the end of the first paragraph in Section 2.

@@

\item \emph{In page 10, line 13, ``We also assume that post-convergence check will usually pass in the first iterations''. I’m not sure it is reasonable.}

  We now provide both a citation and our own investigations (in the appendix) to support this assumption.

\item \emph{In page 8, line 19, ``the reference solutions are continually updated''. The reference solutions are not always updated. So, I’m not sure the term ``continually'' is appropriate.}

  Very good point. In the revised manuscript, we have changed this to ``the reference gets updated systematically'', which is a more accurate description.

@@

\item \emph{In Figure 3, the green line (HSSR) pattern is quite different with others. Could you explain it?}

    We add a brief explanation that HSSR works better when $\lambda$ values in the path are closer to $\lambda_{\max}$ but we decide not to go deeper as HSSR is not our main interest.

@@

\item \emph{In Section 6(Conclusion), there does not exist the sequential safe rules for sparse Cox regression, Poisson regression, and support vector machines. It would be good to explain the existence of the sequential strong rules for each problem.}

    Thanks for this good suggestion. We explain their existence in the last sentence in this section.

@@

\item \emph{In Theorem 4.1, $x_* \in \{ x_j : \hat{\beta}_j(\lambda_l) \ne 0 \}$. Is the $x_*$ arbitrary? It seems that further explanation for $x_*$ is needed.}

    It is a good question. In the second sentence in the last paragraph after the theorem, we clarify that $x_*$ can be chosen arbitrarily and add more explanation for it.

@@

\textbf{Minor corrections:}

\begin{enumerate}

\item \emph{In Theorem 4.1, $\lambda_0 := \max_j |x^T_j y/n|$ may be incorrect. It would be the maximum gradient value at $\hat{\beta} = 0$ of the negative log-likelihood in logistic regression model.}

    Under the standardization assumption, the gradient will be $|x^T_j y/n|$ and this expression is correct.

@@

\item \emph{In page 10, line 20, $|S_{l=b}|$ should be $|S_{l+b}|$.}

    Thanks for the good observation and it is fixed.

\item \emph{In Figure 1, the green line is not continuous. It would be better to modify Figure 1.}

    The green line should have jumps that represent updates of reference that increase screening power. We add an explanation of the jumps in the 7th sentence in the paragraph after the figure.

### 2021-09-09 jscs/adaptive-hssr.tex
Fix appendix and add responses

>>>>

\section*{Appendix}
%\label{sec:ap}

To illustrate the inner workings of the proposed algorithm on a variety of problems, for a single replication of each of the four data sets in Section~\ref{sec:real-data} we recorded the number of features discarded by each screening method at each $\lambda$ value along the path of $L=100$ values. Features discarded by SEDPP are safely discarded, while features discarded by SSR are not necessarily safe and require additional checks. For the adaptive hybrid rule, we use AHR\_Safe to denote the features that are safely discarded first by adaptive EDPP and AHR\_Strong to denote features that discarded afterward by adaptive SSR and require additional checks. The results are summarized in Figure \ref{fig:ap1}.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.82\textwidth]{app1.eps}    \caption{Comparing number of discards by screening methods for lasso model along $\lambda$ values.}
    \label{fig:ap1}
\end{figure}

Note that the AHR method is rather ``jumpy''; this is by design, and these spikes represent moments along the path that triggered an update of the reference solution to gain more screening power. Our adaptive updating algorithm determines the timing of update dynamically depending on the screening performance. For all four data sets, with about only 10 updates, the adaptive hybrid screening can safely discard a great number of features at most $\lambda$ values.

Next we compare the number of rounds of KKT condition checks and the number of KKT condition checks on individual features needed for AHR and SSR respectively, as shown in Table \ref{Tab:kkt1} and Table \ref{Tab:kkt2}. In a round of KKT condition checking, all features that are not discarded by the safe rule are checked, which potentially includes a large number of individual feature checks.

\begin{table}[H]
\centering
\begin{tabular}{lllll}
\toprule
Screening method & GENE & MNIST & GWAS & NYT \\
\midrule
SSR & 99 & 99 & 99 & 99 \\
AHR & 99 & 99 & 99 & 99 \\
\bottomrule
\end{tabular}
\caption{Number of rounds of KKT conditions checks required for the whole path}
\label{Tab:kkt1}
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{lllll}
\toprule
Screening method & GENE & MNIST & GWAS & NYT \\
\midrule
SSR & 1,712 & 5,943 & 60,986 & 5,429 \\
AHR & 172 & 286 & 2,353 & 622 \\
\bottomrule
\end{tabular}
\caption{Number of KKT conditions checks on individual features ($\times10^3$) required for the whole path}
\label{Tab:kkt2}
\end{table}

Although all paths includes $L=100$ $\lambda$ values, the solution at the first value is exactly $\hat{\beta}(\lambda_0)=0$ and does not require any checking. Thus, if no violations occur in the application of the strong rule, the number of rounds of KKT condition checking required will attain its minimum value of $99$. In all four data sets, neither SSR nor the adaptive SSR in AHR is ever violated (as other authors have discussed, violations are rare but possible). However, the two methods are very different in terms of the number of individual features involved in these condition checks. AHR requires far fewer checks on individual features because the adaptive EDPP narrows down the set of features that need to be checked by an order of magnitude, which is how it achieves greater computational efficiency.
